---
title: "Klamath_Meadows_Report"
date: "Updated: `r format(Sys.Date())`"
output:
  html_document:
    highlight: pygments
    theme: yeti
    code_folding: hide
    fig_caption: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)

suppressPackageStartupMessages({
  library(tidyverse);
  library(here);
  library(lubridate);
  library(viridis);
  library(magrittr);
  library(sf);
  library(ggmap);
  library(ggthemes); # theme_map()
  library(sp);
  library(raster);
  library(mapview);
  library(maptools); 
  library(RColorBrewer);
  library(ggrepel);
  library(cowplot);
  library(kableExtra)
})
```


# Klamath Meadow Aerial Mapping Summary

Remote-sensing mapping of meadows in the Klamath USFS Northern Province was conducted in 2020. Layers used to delineate and digitize meadow polygons included:

 - HUC 10 and HUC12 watershed boundaries
 - USFS property boundaries
 - ESRI hi-res aerial imagery basemap (imagery dates vary across extent)
 - Sentinel2 10m resolution imagery from other times of year
 - NDVI greenest pixel compilation from LANDSAT and MODIS imagery (generated using Google Earth Engine)
 - National Wetlands Inventory (NWI) polygons
 - National Hydrography Dataset (NHD+ streamlines) to check for stream presence


```{r getmdws}

db <- paste0(here(),"/data/mdw_spatial_data.gpkg")
st_layers(db)
h10 <- read_sf(dsn = db, layer="huc10_focus", quiet=T)
h12 <- read_sf(dsn = db, layer="huc12_focus", quiet=T) 
nhd_springs <- read_sf(dsn = db, layer="nhd_springs_focus")
usfs_focus <- read_sf(dsn = db, layer="usfs_focus")
ownership_focus <- read_sf(dsn = db, layer="ownership_focus")
mdws <- read_sf(dsn=db, layer="mdws_klam_digitized")

```

## Overview Map

The following shows an interactive map with all the meadows delineated by remote-sensing methods. Only USDA Forest Service properties were delineated, but this effort may be expanded following similar methods.

```{r meadow map}
mapbases <- c("Esri.WorldTopoMap","Esri.WorldGrayCanvas", "Esri.WorldImagery", 
              "Stamen.TonerLite","OpenTopoMap", "CartoDB.PositronNoLabels", 
              "OpenStreetMap")

# mdw map
  mapview(usfs_focus, map.types=mapbases, col.regions=NA, color="brown", lwd=2, 
          alpha.regions=0, legend=F) + 
  mapview(ownership_focus, map.types=mapbases, zcol="OWNERCLASS",  lwd=.7, 
          alpha.regions=0.1, layer.name="Land Owner") +
  mapview(h10, map.types=mapbases, col.regions=NA, color="blue", lwd=2, 
        alpha.regions=0, legend=F) + 
  mapview(h12, map.types=mapbases, col.regions=NA, color="skyblue", lwd=1, 
          alpha.regions=0, legend=F) +
  mapview(mdws, map.types=mapbases, color="seagreen", col.regions="seagreen2", 
        lwd=2, alpha.regions=0.1, legend=TRUE, layer.name="Meadows")
```

## Meadow Delineation Results

```{r}

# make a dataframe and drop sf geometry
mdws_df <- st_drop_geometry(mdws)

# replacy Y/N with 1/0
mdws_df <- mdws_df %>% 
  mutate_if(is.character, list(~ifelse(.=="Y", 1, .)))
mdws_df <- mdws_df %>% 
  mutate_if(is.character, list(~ifelse(.== "N", 0, .)))
# convert to numeric
mdws_df <- mdws_df %>% 
  mutate_at(vars("STREAM_YN","TREES_YN","TREES_G25","ROAD_YN"),as.integer)

# drop cols we aren't interested in for summary
mdws_df <- mdws_df %>% 
  mutate(NWI_FID=ifelse(NWI_FID==0, " ", NWI_FID), 
         NOTES=ifelse(is.na(NOTES), " ", NOTES))

# summarize
mdws_summ <- mdws_df %>% 
  summarize_at(vars("STREAM_YN","TREES_YN","TREES_G25","ROAD_YN", "AREA_HA", "AREA_KM", "AREA_AC"), sum, na.rm=TRUE) %>% 
  mutate(mdw_count = nrow(mdws_df)) %>% 
  dplyr::select(mdw_count, everything()) %>% 
  set_names(c("Total Meadows", "Contain Streams", "Contain Trees", "Contain Roads", "Contain Tree Cover > 25%", "Total Area (ha)",
                      "Total Area (km2)", "Total Area (acres)"))

kable(mdws_summ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

mdws_df$CONFIDENCE <- factor(mdws_df$CONFIDENCE, levels = c("High", "Moderate", "Low"))
mdws_conf <- data.frame(table(mdws_df$CONFIDENCE))
names(mdws_conf) <- c("Confidence", "Number of Meadows")
kable(mdws_conf) %>%
  kable_styling(full_width = F, position = "left", 
                bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r}
mdws_df <- mdws_df %>% 
  mutate(AREA_HA = round(AREA_HA, 1), 
         AREA_KM = round(AREA_KM, 3), 
         AREA_AC = (round(AREA_AC,1))) %>% 
  dplyr::select(`FS Block` = BLOCK_ID, 
                `Contains Stream` = STREAM_YN, 
                `Contains Trees` = TREES_YN,
                `Tree Cover > 25%` = TREES_G25, 
                Road = ROAD_YN, 
                `Area (ha)` = AREA_HA, 
                `Area (acre)` = AREA_AC,
                `Perimeter (m)` = PERIMETER_M, 
                `Average Elevation` = ELEV_MEAN_M, 
                `Edge Complexity` = EDGE_COMPLEXITY,
                `NWI Poly FID` = NWI_FID, 
                Confidence = CONFIDENCE, Notes = NOTES)
# make into table
kable(mdws_df) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  scroll_box(height = "400px")
```


### High Confidence

```{r, echo=F, out.width = 1200, fig.cap = 'Figure 1. Examples of meadows designated "high confidence".'}

knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_high-confidence.png"))

```

### Moderate Confidence

```{r, echo=F, out.width = 1200, fig.cap = 'Figure 2. Examples of meadows designated “moderate confidence".'}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_mod-confidence.png"))
```

### Low Confidence

```{r, echo=F, out.width = 1200, fig.cap = 'Figure 3. Examples of meadows designated “low confidence".'}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_low-confidence.png"))
```

### Non-Meadow Examples

```{r, echo=F, out.width = 1200, fig.cap = 'Figure 4. Examples of open areas not considered meadows, such as areas comprised of scrub and bare ground (L) or regrowing clearcuts (R).'}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_not-meadows.png"))
```



# Delineation Challenges and Limitations

### Boundary Questions

```{r, echo=F, out.width = 1200, fig.cap = "Figure 5. Examples of uncertain boundaries: some meadows appear to be hydrologically connected through forested areas (L), and the boundaries between dry grass, scrub, and bare ground are often unclear in satellite imagery (R)."}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_questionable-border1.png"))
```


```{r, echo=F, out.width = 1200, fig.cap = "Figure 6. Examples of open wetland areas that are borderline wetland cases. Should wetlands with complete shrub coverate (L) be included, or should the dataset be restricted to wetlands with at least some open areas? Should wetter streamside areas that are shrub-dominated (R) compared to drier, scrubby upland areas be included?"}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_questionable-border2.png"))
```



### Imagery Date Matters

```{r, echo=F, out.width = 1200, fig.cap = "Figure 7. Importance of time of year: meadows look very different early (L) and late (R) in the growing season."}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_seasonality.png"))
```


```{r, echo=F, out.width = 1200, fig.cap = "Figure 8. Utility of (L) NDVI greenest pixel and (R) sentinel imagery from different time of year in distinguishing bare ground from dry grass in late-season high-res imagery (C)."}
knitr::include_graphics(path=paste0(here::here(),"/figs/klam-mdw-examples_seasonality-NDVI-sentinel.png"))
```

Calculating NDVI (an index of vegetation greenness/vigor) throughout the year and creating a "greenest pixel" composite (left image) is also useful in distinguishing vegetation from bare ground; variation in maximum greenness can also help distinguish between areas that have vigorously growing vegetation during part of the year (e.g. emergent vegetation or early-successional, post-clearcut areas) and patches of evergreen trees or shrubs. Publicly available, lower-resolution imagery from early in the season (right image) can be useful in differentiating dry, late-season emergent vegetation and bare ground when high-resolution imagery is from late in the year (center image); when high-resolution imagery is from early in the year, later-season imagery may help differentiate between shrub/scrub and emergent vegetation.
